---
title: "Seagrass Fluxes - Worked Example"
author: "Jon Lefcheck"
date: "February 20, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = getwd())

# load required libraries
library(fluxweb)
library(nlme)
library(readxl)
library(piecewiseSEM)
library(tidyverse)

# define years
years <- c(2007:2013)
```

This is a worked example of how to compute energy fluxes through seagrass food webs in R. The data come from a long-term survey of seagrass communities at Goodwin Island in the York River Estuary, Chesapeake Bay, USA monthly from 2007-2013. Sampling consisted of two transects laid parallel to shore (one inshore and one offshore). Five samples were randomly taken along each transect for a total _n_ = 10 per month. More details can be found in: Douglass, J. G., Douglass, J. G., France, K. E., France, K. E., Richardson, J. P., & Duffy, J. E. (2010). Seasonal and interannual change in a Chesapeake Bay eelgrass community: insights into biotic and abiotic control of community structure. Limnology and Oceanography, 55(4), 1499-1520.

## Data Preparation

First, we will briefly explain and clean up the data.

Primary producers were sampled in several ways. First, a 15-cm diameter biomass core was taken at each replicate to quantify above- and belowground biomasses of seagrasses (two species: _Zostera marina_ and _Ruppia maritima_), which also included drift macroalgae. Samples were taken back to the lab and dried to constant weight (in g). These values were further scaled to 1 m^2 of bottom area.

```{r}
# load biomass core data
grass <- read.csv("./Data/grass.csv")


# scale grass/macroalgal dry weight to 1 m^2 bottom area
grass$dm <- grass$dm * (1 / (pi * (0.15/2)^2))
```

Second, a single shoot of _Zostera marina_ was collected at each replicate, brought back to the lab and scraped of epiphytes. The epiphytes were filtered, extracted in 90% acetone, and then absorbances measured using a spectrophotometer, converted to biomass using the trichromatic equations of Lorenzen 1967, and normalized to leaf area. 

```{r}
# load epiphyte data
epi <- read.csv("./Data/epi.csv")


# compute epiphytic chl-a in ug using trichromatic equations
epi$date <- as.character(epi$date)

colnames(epi) <- gsub("X", "", colnames(epi))

epi <- do.call(rbind, lapply(unique(epi$date), function(i) {
  
  x <- subset(epi, date == i)
  
  # subtract average of blanks
  if("Blank" %in% x$replicate) {
    
    x[x$replicate != "Blank", grepl("nm", colnames(x))] <- x[x$replicate != "Blank", grepl("nm", colnames(x))] - 
      colMeans(x[x$replicate == "Blank", grepl("nm", colnames(x))], na.rm = T)
    
    x <- subset(x, replicate != "Blank")
    
  }
  
  # set negative absorbances to 0
  x[, grepl("nm", colnames(x))][x[, grepl("nm", colnames(x))] < 0] <- 0
  
  # use trichromatic equations to convert absorbance into ug mL-1
  x$chla <- 11.85 * (x$`664nm` - x$`750nm`) - 1.54 * (x$`647nm` - x$`750nm`) - 0.08 * (x$`630nm` - x$`750nm`)
  
  x$chlb <- 21.03 * (x$`647nm` - x$`750nm`) - 5.43 * (x$`664nm` - x$`750nm`) - 2.66 * (x$`630nm` - x$`750nm`)
  
  x$chlc <- 24.52 * (x$`630nm` - x$`750nm`) - 1.67 * (x$`664nm` - x$`750nm`) - 7.60 * (x$`647nm` - x$`750nm`)
  
  x$chlp <- 7.60 * ((x$`480nm` - 3 * x$`750nm`) - 1.49 * (x$`510nm` - 2 * x$`750nm`))
  
  
  # multiply by volume of sample (20 mL) to yield ug biomass
  x[, grepl("chl", colnames(x))] <- x[, grepl("chl", colnames(x))] * 20
  
  # convert leaf area to dry weight
  leaf.dw <- rowMeans( -6.921e-5 + 2.27e-3 * x[, grepl("Area", colnames(x))], na.rm = T)
  
  chla.g.dw <- (x$chla * 1e-6) / leaf.dw 
  
  data.frame(
    x[, c("year", "month", "day", "location", "inshore.offshore", "replicate")],
    chla.g.dw
  )
  
} ) )
```

Alternately, epiphytes could be scraped, dried, and weighed to obtain an estimate of fouling biomass.

Epiphytes were further scaled to the dry weight of seagrass in the corresponding biomass core to express epiphytic algal in units of g per m^2. The seagrasses were then removed from the dataset, as nothing consumes them in this part of the world.

```{r}
# scale epi to dry weight zostera
grass.z <- grass %>% filter(species.name.revised == "Zostera marina blades")

epi <- left_join(epi, grass.z)

epi$total.chla.g <- epi$chla.g.dw * epi$dm

epi$species.name.revised <- "epiphytic.algae"

# remove seagrasses from grass as nothing eats them
grass <- filter(grass, !species.name.revised %in% c("Zostera marina blades", "Ruppia maritima blades"))
```

Next, we sampled epifaunal invertebrates using a grab sample. Samples were taken back to the lab, and associated macrophytes dried and weighed. Invertebrates were sorted to species, then passed through a series of nested sieves to obtain size-fractionated abundance. Biomass was estimated from these values using the equations in Edgar (1991), normalized to macrophyte biomass, and then expressed per m^2 of bottom area.

```{r}
bugs <- read.csv("./Data/bugs.csv")

# get function to apply Edgar equations
source("https://gist.githubusercontent.com/jslefche/c480eeb2ad67ca7c9a46/raw/177e47862e4053337cca66e94f5f0b91d5e7086f/edgarMethod.R")

# compute bug biomass from size fractionated abundances
bugs <- filter(bugs, group %in% c("Crustacean", "Mollusc", "Polychaete", "Caprellid"))

bugs <- edgarMethod(bugs, group = "group")

bugs$total.biomass.g <- bugs$total.biomass / 1000

# scale abundances/biomasses to 1 m^2 bottom area
bugs[, grepl("abund|biomass", colnames(bugs))] <- bugs[, grepl("abund|biomass", colnames(bugs))] * (1 / (0.2 * 0.2))
```

Mesopredators were censused using a 5-m dip-net sweep perpendicular to the main transect. All organisms were measured (total length for fishes, carapace width for crabs) and biomass was estimated using length-weight regressions from FishBase or Fredette & Diaz (1990). The final values were then expressed per m^2 of bottom area.

```{r}
preds <- read.csv("./Data/preds.csv")

# compute predator biomass from L-W regressions
predB <- function(x) {
  
  a <- c(`Syngnathus sp.` = 0.00038, `Callinectes sapidus` = 0.0643, 
         `Paralichthys dentatus` = 0.01, `Palaemonetes sp.` = 0.588,
         `Syngnathus fuscus` = 0.00038, `Syngnathus floridae` = 0.0004,
         `Crangon septemspinosa` = 0.5999, `Leiostomus xanthurus` = 0.00871,
         `Gobiosoma bosc` = 0.00617, `Gobiesox strumosus` = 0.0128,
         `Symphurus plagiusa` = 0.00794 , `Opsanus tau` = 0.00955,
         `Fundulus heteroclitus` = 0.01072, `Lucania parva` = 0.02188,
         `Trinectes maculatus` = 0.01259, `Pomatomus saltatrix` = 0.01,
         `Cynoscion nebulosus` = 0.0131)

  b <- c(`Syngnathus sp.` = 3.13, `Callinectes sapidus` = 2.74, 
         `Paralichthys dentatus` = 3.16, `Palaemonetes sp.` = 2.53,
         `Syngnathus fuscus` = 3.13, `Syngnathus floridae` = 3.12,
         `Crangon septemspinosa` = 2.41, `Leiostomus xanthurus` = 3.08,
         `Gobiosoma bosc` = 3.08, `Gobiesox strumosus` = 3.03600,
         `Symphurus plagiusa` = 3.02, `Opsanus tau` = 3.07,
         `Fundulus heteroclitus` = 3.17, `Lucania parva` = 3.22,
         `Trinectes maculatus` = 3.08, `Pomatomus saltatrix` = 2.98,
         `Cynoscion nebulosus` = 3)
  
  x$total.biomass.g <- a[x$species.name.revised] * x$size ^ b[x$species.name.revised]
  
  return(x)
  
}

preds <- predB(preds)

# remove species without biomass
preds <- preds %>% filter(!is.na(total.biomass.g))

# scale predator abundance/biomass to 1 m^2 bottom area
preds[, grepl("abundance|biomass", colnames(preds))] <- preds[, grepl("abundance|biomass", colnames(preds))] * (1 / (0.52 * 5))
```

Finally, abiotic parameters were measured using an in situ data sonde maintained by the Chesapeake Bay NERR and located within the seagrass bed. Values are daily averages.

```{r}
env <- read.csv("./Data/env.csv")
```

## Quantifying Trophic Relationships

The framework from Barnes et al. requires knowledge of basic trophic relationships. These can be extremely general (e.g., herbivores, carnivores) or highly-resolved (e.g., detritivorous amphipods, omnivorous fishes). For this example, we will draw on a priori knowledge of the trophic roles of organisms to construct a very rudimentary food web (primary producers -> herbivores -> predators). This effort could be further resolved using relationships derived from the literature, or by running targeted experiments or applying stable isotope analysis.

The final matrix consists of an _s_ by _s_ matrix of _s_ species, with a 1 if the species in column _j_ consumes the organisms in row _i_ and a 0 if they do not interact.

```{r}
# get vector of primary producer names
pp.names <- c(as.character(unique(grass$species.name.revised)), "epiphytic.algae")

# get vector of predator names
pred.names <- c("Nereis sp.", "Palaemonetes vulgaris", "Palaemonetes sp.", as.character(unique(preds$species.name.revised)))

# get vectors of species
species <- c(as.character(unique(bugs$species.name.revised)), as.character(unique(preds$species.name.revised)),
             as.character(unique(grass$species.name.revised)), as.character(unique(epi$species.name.revised)))

# get vector of herbivore names
herb.names <- species[!species %in% c(pp.names, pred.names)]

# create diet matrix
mat <- matrix(0, nrow = length(species), ncol = length(species))

dimnames(mat) <- list(species, species)

# herbivores eat algae
mat[rownames(mat) %in% pp.names, colnames(mat) %in% herb.names] <- 1

# preds eat herbivores
mat[rownames(mat) %in% herb.names, colnames(mat) %in% pred.names] <- 1
```

## Estimating Fluxes

The framework assumes that some energy is lost through both respiration (metabolism) and reduced assimilation efficiency.

Population-specific metabolic rates are estimated using average body size and eqn. 2 in Brown et al. (2004): $$I = I_0M^{b}$$

The intercept ($$I_0$$) and exponent $$b$$ can be refined based on whether the organism is an inverberate or ectothermic (fishes) (See Table 1, Gauzens et al. (2018)).


```{r}
# get vector of average biomasses to scale losses (per g  biomass)
indiv.bugs <- cbind.data.frame(species.name.revised = bugs$species.name.revised, avg = rowMeans(bugs[, 24:32]/bugs[, 11:19], na.rm = T)/1000) # convert to g from mg

indiv.bugs <- indiv.bugs %>% group_by(species.name.revised) %>% summarize(avg = mean(avg, na.rm = T))

indiv.preds <- preds %>% group_by(species.name.revised) %>% summarize(avg = mean(total.biomass.g / abundance))

# NAs for primary producers
indiv.grass <- data.frame(avg = rep(0, length(unique(grass$species.name.revised)))) 

indiv.epi <- data.frame(avg = rep(0, length(unique(epi$species.name.revised))))

bodymasses <- c(indiv.bugs$avg, indiv.preds$avg, indiv.grass$avg, indiv.epi$avg)

names(bodymasses) <- c(as.character(indiv.bugs$species.name.revised), as.character(indiv.preds$species.name.revised), 
                       as.character(unique(grass$species.name.revised)), "epiphytic.algae")

# compute metabolic rates for inverts/ectotherms
constants <- data.frame(
  species,
  a = ifelse(species %in% c(herb.names, pred.names[c(1:4)]), 17.17, ifelse(species %in% pred.names[-c(1:3)], 18.18, 0)),
  b = ifelse(species %in% pp.names, 0, -0.29)
)

losses <- constants[match(names(bodymasses), constants$species), "a"] * bodymasses ^
  constants[match(names(bodymasses), constants$species), "b"]
```

This equation could further incorporate the effects of temperature on metabolic rates (eqn. 4, Brown et al. (2004)), but for purposes of simplicity, we will proceed with the estimates derived from body size alone.

Similarly, feeding efficiences can be estimated from literature values (see Lang et al. (2017)) based on whether the organisms are plants, animals, or detritus. 

```{r}
# compute vector of assimilation efficiencies
org.types <- ifelse(names(losses) %in% pp.names, "plant", "animal")

eff <- c(animal = 0.906, plant = 0.545, detritus = 0.158)

efficiences <- eff[org.types]

names(efficiences) <- names(losses)
```

We now have all the information necessary to compute fluxes using the `fluxweb` package. In this case, we will compute the total biomass across all 10 replicates (which is more representative of the whole food web), and divide by 10 (or the actual number of replicates, if any were lost) to obtain the flux per m^2. We will also weight the prey preferences by total biomass of their prey to account for month-to-month differences in population biomasses.

```{r}
# compute fluxes for each month
fluxes <- do.call(rbind, lapply(years, function(i) {
  
  do.call(rbind, lapply(1:12, function(j) {
  
    bugs.sub <- bugs %>% filter(year == i & month == j)
    
    preds.sub <- preds %>% filter(year == i & month == j)
    
    grass.sub <- grass %>% filter(year == i & month == j)
    
    epi.sub <- epi %>% filter(year == i & month == j)
    
    if(nrow(bugs.sub) == 0 | nrow(preds.sub) == 0 | nrow(epi.sub) == 0 | all(is.na(epi.sub$total.chla.g)))
      
      data.frame() else {
    
    # a vector with the total biomass of each population (per m^2 of bottom area)
    total.bugs <- bugs.sub %>% group_by(species.name.revised) %>% summarize(sm = sum(total.biomass.g))
    
    total.preds <- preds.sub %>% group_by(species.name.revised) %>% summarize(sm = sum(total.biomass.g))
    
    total.grass <- grass.sub %>% group_by(species.name.revised) %>% summarize(sm = sum(dm))
    
    total.epi <- epi.sub %>% summarize(sm = sum(total.chla.g, na.rm = T)) %>% cbind.data.frame(species.name.revised = "epiphytic.algae", .)
    
    biomasses.df <- rbind(total.bugs, total.preds, total.grass, total.epi)
    
    biomasses.df <- filter(biomasses.df, !is.na(species.name.revised) | sm != 0)
    
    biomasses <- biomasses.df$sm
    
    names(biomasses) <- biomasses.df$species.name.revised
    
    biomasses[biomasses < 0] <- 0
  
    # subset efficiencies
    efficiences.sub <- efficiences[names(biomasses)]
    
    # subset interaction matrix
    mat.sub <- mat[names(biomasses), names(biomasses)]
    
    # subset bodymasses
    losses.sub <- losses[names(biomasses)]
    
    # compute fluxes
    mat.fluxes <- fluxing(mat.sub, biomasses, losses.sub, efficiences.sub)
    
    # get fluxes to herbivores
    herbivory <- sum(rowSums(mat.fluxes[rownames(mat.fluxes) %in% pp.names, , drop = F]))
    
    carnivory <- sum(rowSums(mat.fluxes[rownames(mat.fluxes) %in% herb.names, , drop = F]))
    
    data.frame(
      year = i,
      month = j,
      flux = c("Herbivory", "Predation"),
      value = c(herbivory, carnivory) / nrow(unique(epi.sub[, c("inshore.offshore", "replicate")])) / 12
    )
    
    }
    
  } ) )
  
} ) )

```

## Plotting Results

Now, let's plot the trends through the months of the year:
```{r}
# plot results
fluxes.summary <- fluxes %>% group_by(month, flux) %>% 
  
  summarize(mean.value = mean(value, na.rm = T), se.value = plotrix::std.error(value, na.rm = T))

fluxes.summary$flux <- factor(fluxes.summary$flux, levels = c("Herbivory", "Predation"))

ggplot(fluxes.summary, aes(x = as.factor(month), y = mean.value)) +
  geom_line(aes(group = 1)) +
  geom_errorbar(aes(ymax = mean.value + se.value, ymin = mean.value - se.value), width = 0) +
  geom_point() +
  facet_grid(~ flux) +
  labs(x = "Month", y = bquote(Fluxes~(Joules~30~d^-1~m^-2))) +
  theme_bw(base_size = 18) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
```

We see that fluxes are quite low in the colder winter months (March, April, November) but increase rapidly in the spring though mid-summer, when production is highest. There is a dip in August, when eelgrass naturally senesces in this part of the world. There is a slight bump in the early fall due to the slight resurgence in seagrass during this period, but by then much of the faunal community has emigrated from the Bay. Notably, fluxes are lower for predators than for herbivores.

We can further relate these values to the biotic and abiotic properties of the ecosystem. For example, how does the richness of species in each month influence fluxes? Theory suggests that more diverse communities are more efficient at using available resources, translating to higher energy flux.

```{r}
# plot against diversity
bugs.S <- bugs %>% group_by(year, month) %>% summarize(richness = length(unique(species.name.revised))) %>%
  
  mutate(flux = "Herbivory") 

fluxes <- preds %>% group_by(year, month) %>% summarize(richness = length(unique(species.name.revised))) %>%
  
  mutate(flux = "Predation") %>% rbind(., bugs.S) %>% left_join(fluxes, .)

ggplot() +
  geom_point(data = fluxes, aes(x = log10(richness), y = log10(value), group = flux, col = flux)) +
  stat_smooth(method = "lm", data = fluxes, aes(x = log10(richness), y = log10(value), group = flux, col = flux), 
              se = F) +
  scale_color_manual(values = c("red", "blue"), name = "") +
  labs(x = bquote(log[10]~(Species~Richness)), y = bquote(log[10]~(J~30~d^-1~m^-2))) +
  theme_bw(base_size = 18) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = c(0.6, 0.2)
  )
```

Here, we see that predator richness enhances fluxes, but they are largely unaffected by changes in herbivore diversity. This supports the now well-described increase in diversity effects with increasing trophic level (Lefcheck et al. (2015), O'Connor et al. (2017)).

Similarly, metabolic theory indicates that higher temperatures should drive greater fluxes due to higher metabolic demands.

```{r}
env$month <- as.numeric(env$month)

fluxes <- env %>% filter(month %in% unique(fluxes$month)) %>% group_by(month) %>% 
  
  summarize(temp = mean(Temp, na.rm = T), sal = mean(Sal, na.rm = T), turb = mean(Turb, na.rm = T)) %>%
  
  left_join(fluxes, .)

ggplot() +
  geom_point(data = fluxes, aes(x = temp, y = log10(value), group = flux, col = flux)) +
  stat_smooth(method = "lm", data = fluxes, aes(x = temp, y = log10(value), group = flux, col = flux), 
              se = F) +
  scale_color_manual(values = c("red", "blue"), name = "") +
  labs(x = bquote(Temp~degree~C), y = bquote(log[10]~(Joules~30~d^-1~m^-2))) +
  theme_bw(base_size = 18) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = c(0.6, 0.2)
  )
```

Indeed, as temperature increases, the estimated fluxes to both herbivores and predators increases (although non-linearly).

We can use linear models to test the effects of each of these properties while controlling for the others. This will tell us whether biodiversity is still important given the influence of temperature.

```{r}
herbivory.mod <- lme(log10(value) ~ log10(richness) + log10(temp) + log10(sal), random = list(year = ~1, month = ~1), data = subset(fluxes, flux == "Herbivory"))

car::vif(herbivory.mod)

coefs(herbivory.mod)

rsquared(herbivory.mod)

predation.mod <- lme(log10(value) ~ log10(richness) + log10(temp) + log10(sal), random = list(year = ~1, month = ~1), data = subset(fluxes, flux == "Predation"))

car::vif(predation.mod)

coefs(predation.mod)

rsquared(predation.mod)

```

These models could be further refined to show the indirect effect of temperature on biodiversity using techniques like structural equation modeling.

